### 2015-11-05 wave_26oct/presentation/methods_results_5_nov.Rmd
Analyze data from pilot study

---
title: "Methods Results 5 November"
author: "Daniel Wall"
date: "November 5, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
setwd("/git_repositories/timeratios/wave_26oct/")
source("load_packages.R")
rm(list = ls())
load("analyzed_data/analyzed_data.Rdata")
```

# Summary

After cleaning the data we had a total of `r n_distinct(pzwhl$responseid)` participants.

## Items that the participants saw:

Note: These are measured on a 0-100 scale.

### Small ratio change

![alt text](small_ratio_change.png)

### Large ratio change

![alt text](big_ratio_change.png)

# Results

## Multilevel model predicting perceived dissimilarity between the items

I ran a linear multilevel model with varying intercepts for subject. Main effects of ratio dissimilarity type (small or large) `rs` and `rb` in the model, item (adding 0, 10, or 20 days to the initial amount) `item1`, `item2`, `item3` in the model, and their interaction. All variables are dummy coded

### Anova with type 3 sums of squares

```{r, echo=FALSE}
Anova(diss_lmer, type = 3)
```

We see a main effect of ratio dissimilarity type, item, and their interaction.

### Model Summary

```{r, echo=FALSE}
summary(diss_lmer)
```

`diss_typers` is the estimate of large ratio dissimilarity minus the estimate of small ratio dissimilarity, while linearly adjusting for all other effects.

`item2` is the estimate of the 0 day change in time minus the 10 day increase in time, while linearly adjusting for all other effects.

### interaction graph for zero type and variance type

The x axis is the perceived dissimilarity between the measures.

```{r, echo=FALSE}

ggplot(eff_df, aes(x = item, y = fit, group = diss_type, color = diss_type)) +
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper))

```

We see no significant change in percieved dissimilarity by adding time to both delays when ratios are small, but a decrease in percieved dissimilarity by adding time to both delays when ratios are large.