### 2015-11-13 shinyapps/scholtenReadSanborn2014.Rmd
Create Shiny app for SRS

---
title: "Scholten Read Sanborn"
author: "Daniel Wall"
date: "November 12, 2015"
output: html_document
runtime: shiny
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(shinyapps)
library(dplyr)
```

```{r, echo=FALSE}
prob_dist <- function (p, alpha) {
  exp(-(-log(p))^alpha)
}

prob_df <- data.frame(prob = seq(0, 1, .01))

```

## Exponential Discounting

In the exponential discounting model, the decision maker will be indifferent between SS and LL when

$$\delta^{t_S}x_S = \delta^{t_L}x_L$$

Where $0 < \delta < 1$ is a one period discount factor and $\rho = 1 - \delta$

One period discount factor, an empiricial measure of discounting over that interval:

$$\delta = (\frac{x_S}{x_L})^{\frac{1}{t_L - t_S}}$$

how does the on period discount factor change for different values  of smaller sooner amount/time, and larger later amount/time.

```{r, echo=FALSE}
exp_disfac <- function(ssamt, sstime, llamt, lltime) {
  (ssamt/llamt)^(1/(lltime - sstime))
}

```

```{r, echo=FALSE}
inputPanel(
  sliderInput("ssamt", "Smaller Sooner $", min=0, max = 500, step=5, value= 10),
    sliderInput("llamt", "Larger Later $", min=0, max = 500, step=5, value= 20),
  sliderInput("sstime", "Smaller Sooner time", min=0, max = 365, step=1, value= 1),
    sliderInput("lltime", "Larger Later time", min=0, max = 365, step=1, value= 2)

  )

renderPrint({
exp_disfac(input$ssamt, input$sstime, input$llamt, input$lltime)

})

```

# Hyperbolic

According to Loewenstein and Prelec’s (1992) hyperbolic discounting model, the deci-
sion maker will be indifferent between SS and LL when

$$d(t_S)v(x_S) = d(t_L)v(x_L)$$

where v is a reference-dependent value function, and d is a generalized hyperbolic dis- count function:

$$d(t) = (1 + \tau t)^{-\beta/\tau}$$

where $\beta > 0$ is delay discounting, and $\tau > 0$ is the departure from exponential discounting. As $\tau$ goes to 0, the generalized hyperbolic discount function reduces to the exponential discount function $d(t)\exp^{-\beta t} = \delta^t$. The hyperbolic discounting model predicts the common difference effect. To see this, the total discounting over the interval tS ? tL can be expressed in the form of a discount **fraction**.

__So discount fraction is different than discount factor__

$$\frac{v(x_S)}{v(x_L)} = (\frac{1+\tau t_S}{1+\tau t_L}) ^{\beta/\tau} = (\frac{1+\tau t_S}{1 + \tau t_S + \tau(t_L-t_x)}) ^{\beta /\tau}$$

```{r}
hyp_disfrac <- function(sstime, lltime, beta, tau) {
  sstime <- as.numeric(sstime)
  lltime <- as.numeric(lltime)
  beta <- as.numeric(beta)
  tau <- as.numeric(tau)

  ((1 + tau * sstime)/(1 + tau * sstime + tau *(lltime - sstime))) ^(beta/tau)
}

hyp_disfrac(0, 1, .1, .1)
```

```{r, echo=FALSE}
inputPanel(
  textInput("sstime", "Smaller Sooner time"),
  textInput("lltime", "Larger Later time"),
  textInput("beta", "beta > 0 delay"),
  textInput("tau", "tau > departure from exponential discounting")
  )

renderPrint({
hyp_disfrac(input$sstime, input$lltime, input$beta, input$tau)

})

```

The value function has 5 properties. First three from Prospect theory

1. Reference Dependence
2. Diminishing Sensitivity
3. Loss aversion

2 additional dealing with elasticity of value function or the ratio between the proportional change in value and the proportional change in amount

$$\epsilon = \frac{v' (x)/v(x)}{1/x}$$

4. Increasing elasticity. For constant proportional increases of x > 0 (decreases of x < 0), v(x) increases (decreases) by increasing proportional amounts. Therefore, if outcome magnitude is increased by a common multiplicative constant, the ratio between the values of xS and xL decreases

5. Loss amplification. Elasticity is greater for losses than for gains. That is, decreasing x < 0 by a given proportion yields a greater propor- tional change in value than increasing x > 0 by the same proportion

```{r}
hyp_val <- function(x, gamma, lambda, mu, sigma) {
  # x <- as.numeric(x)
  if (x >= 0) {
    (1 - gamma) * x^(1-gamma) + (mu *gamma * x) ^ gamma
  } else if (x < 0) {
    -lambda*((1-gamma)*(-x) ^(1-gamma) + (mu + sigma)*gamma*(-x)^gamma)
  } 
}

hyp_val <- function(x, gamma, lambda, mu, sigma) {
  # x <- as.numeric(x)
  ifelse(x >= 0,(1 - gamma) * x^(1-gamma) + (mu *gamma * x) ^ gamma, -lambda*((1-gamma)*(-x) ^(1-gamma) + (mu + sigma)*gamma*(-x)^gamma))

}

hyp_val_df <- data.frame(x = -1000:1000)

hyp_val_df %>% rowwise %>% mutate(value = hyp_val(x, 1, 1, 1, 1))
hyp_val(-10, 1, 1, 1, 1)

```

where ½< c < 1 is diminishing absolute sensitivity to outcomes, l > 0 is increasing elasticity, k > 1 is loss aversion, and r > 0 is loss amplification

```{r, echo=FALSE}
inputPanel(
  sliderInput("gamma", "gamma diminishing absolute sensitivity to outcomes",  min=.1, max = .99, step=.001, value= .1),
  sliderInput("lambda", "Lambda loss aversion",  min=1, max = 4, step=.001, value= .1),
  sliderInput("mu", "mu  increasing elasticity",  min=.1, max = .99, step=.001, value= .1),
  sliderInput("sigma", "sigma loss amplification",  min=.1, max = .99, step=.001, value= .1)
  )

renderPlot({
  hyp_val_df$lambda <- input$lambda
  hyp_val_df$sigma <- input$sigma
  hyp_val_df$mu <- input$mu
  hyp_val_df$gamma <- input$gamma

  hyp_val_df <- transform(hyp_val_df, value = hyp_val(x, gamma, lambda, mu, sigma))

  ggplot(hyp_val_df, aes(x = x, y = value, group = 1)) +
    geom_line() +
    geom_abline(intercept = 0, slope=0) +
    geom_vline(xintercept = 0) +
    xlab("Outcome") +
    ylab("Value") +
    xlim(c(-1000, 1000))
})

```